---
title: "Exploratory Data Analysis"
author: "Shanmukh"
date: "April 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ESTIMATES OF LOCATION
 - Mean
    - The sum of all values divided by the number of values.
 - Weighted mean
    - The sum of all values times a weight divided by the sum of the weights.
 - Median
    - The value such that one-half of the data lies above and below.
 - Weighted median
    - The value such that one-half of the sum of the weights lies above and below the sorted data.
 - Trimmed mean
    - The average of all values after dropping a fixed number of extreme values.

# VARIABILITY METRICS

- Variance
  - The sum of squared deviations from the mean divided by n - 1 where n is the number of data values.
- Standard deviation
  - The square root of the variance.

- Mean absolute deviation
  - The mean of the absolute value of the deviations from the mean.
  
- Median absolute deviation from the median
  - The median of the absolute value of the deviations from the median.

- Range
  - The difference between the largest and the smallest value in a data set.

- Interquartile range
  - The difference between the 75th percentile and the 25th percentile.
  
# EXPLORING THE DISTRIBUTION

- Boxplot
  - A plot introduced by Tukey as a quick way to visualize the distribution of data.

- Frequency table
  - A tally of the count of numeric data values that fall into a set of intervals (bins).

- Histogram
  - A plot of the frequency table with the bins on the x-axis and the count (or proportion) on the y-axis.

- Density plot
  - A smoothed version of the histogram, often based on a kernel density estimate.

#  CORRELATION

- Correlation coefficient
  - A metric that measures the extent to which numeric variables are associated with one another (ranges from -1 to +1).

- Correlation matrix
  - A table where the variables are shown on both rows and columns, and the cell values are the correlations between the variables.

- Scatterplot
  - A plot in which the x-axis is the value of one variable, and the y-axis the value of another.
  
  
  
sd() -- Standard deviation
var() -- Variance
IQR() -- interquartile range  75 percentile - 25 percentime points
range()
diff(range())  -- difference in max and min value
mad() -- median absolute deivation from the median

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(readr)
```



```{r, message=FALSE, warning=FALSE}

USEPAdata_017 <- read_csv("https://raw.githubusercontent.com/shapinna/Codeporfolio/master/Data/USEPAdata%20017.csv")

ozone <- USEPAdata_017

mean(ozone$`Valid Day Count`,na.rm = TRUE)

median(ozone$`Valid Day Count`,na.rm = TRUE)

sd(ozone$`Valid Day Count`,na.rm = TRUE)

var(ozone$`Valid Day Count`,na.rm = TRUE)

IQR(ozone$`Valid Day Count`,na.rm = TRUE)

range(ozone$`Valid Day Count`,na.rm = TRUE)

diff(range(ozone$`Valid Day Count`,na.rm = TRUE))

```

```{r}
names(ozone) <- make.names(names(ozone))
View(ozone)
nrow(ozone)
ncol(ozone)
str(ozone)

```

```{r}
head(ozone[, c(6:7, 10)])
tail(ozone[, c(6:7, 10)])
table(ozone$Date.of.Last.Change)

library(dplyr)

select(ozone, State.Name) %>% unique %>% nrow

unique(ozone$State.Name)


```
```{r}
summary(ozone$Sample.Duration)
```

```{r}
library(nycflights13)

View(flights)

head(flights)

str(flights)

table(flights$dep_delay)

mean(flights$dep_delay,na.rm = TRUE)

```

```{r}
library(readr)
income <- read_csv("C:/Users/Amirtha/Downloads/income.csv")

mean(income$M_weekly, na.rm = TRUE)
mean(income$M_weekly,na.rm = TRUE,trim=0.1)

median(income$M_weekly,na.rm = TRUE)


```
```{r}
sd(income$F_workers)
IQR(income$F_workers)
mad(income$F_workers)
```
```{r}
library(dplyr)

income2<-mutate(income,dff=income$M_weekly-income$F_weekly)

View(income2)

income2 %>% glimpse()

sddff<-sd(income2$dff,na.rm = TRUE)

round(sddff)

```
na.rm=TRUE ignores NA values before calculating
```{r}
sd(c(1,5,6,7)) # calculates standard deviation for a vector

help(sd)# provides documentation on sd function

args(sd)  # provides arguments for sd function



```

The mean() function computes the arithmetic mean.

arguments:X, na.rm= FALSE,trim =0

default value exists for na.rm and trim hence not mandatory

trim denotes trimming the start and end values 0.1 denotes 10% trim in starting and ending.

```{r}
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Calculate average number of views
avg_li <- mean(linkedin)
avg_fb <- mean(facebook)


# Inspect avg_li and avg_fb
print(avg_li)
print(avg_fb)


# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Calculate the mean of the sum
avg_sum<- mean(linkedin+facebook)

# Calculate the trimmed mean of the sum
avg_sum_trimmed <- mean(linkedin+facebook,trim=0.2)

# Inspect both new variables
print(avg_sum)
print(avg_sum_trimmed)

# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, NA, 17, 14)
facebook <- c(17, NA, 5, 16, 8, 13, 14)

# Basic average of linkedin
mean(linkedin)

# Advanced average of linkedin
mean(linkedin,na.rm=TRUE)



```
