---
title: "Exploratory Data Analysis"
author: "Shanmukh"
date: "April 20, 2019"
output:
  html_document:
    toc: TRUE
    toc_depth: 3
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ESTIMATES OF LOCATION

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(readr) # to load the files
library(MASS)  # to access built in data set
library(corrplot)  # to access correlation functions
library(DescTools)  # to access MeanAD function Mean absolute deviation
library(frequency)  # for generating frequency table
library(matrixStats) # for calcuating weightedMedian
```


**USEPAdata_017 data set **
```{r, echo=FALSE,message=FALSE,warning=FALSE}
USEPAdata_017 <- read_csv("https://raw.githubusercontent.com/shapinna/Codeporfolio/master/Data/USEPAdata%20017.csv")

head(USEPAdata_017)[,c(1:5)]
```

#### Structure of dataset

```{r}
str(USEPAdata_017)
```


### Mean
    - The sum of all values divided by the number of values.

    - The mean() function computes the arithmetic mean.

    - arguments:X, na.rm= FALSE,trim =0

    - default value exists for na.rm and trim hence not mandatory

    - trim denotes trimming the start and end values 0.1 denotes 10% trim in starting and ending.

```{r}
mean(USEPAdata_017$`Valid Day Count`,na.rm = TRUE) # na.rm ignores NA values

```

### Weighted mean
    - The sum of all values times a weight divided by the sum of the weights.
    
```{r}
weighted.mean(mtcars$mpg,w=mtcars$wt)

```
    
### Median
    - The value such that one-half of the data lies above and below.
    
```{r}

median(USEPAdata_017$`Valid Day Count`,na.rm = TRUE)# na.rm ignores NA values
```
    
### Weighted median
    - The value such that one-half of the sum of the weights lies above and below the sorted data.
    
```{r}
weightedMedian(mtcars$mpg,w=mtcars$wt)
```
    
    
### Trimmed mean
    - The average of all values after dropping a fixed number of extreme values.

```{r}
mean(USEPAdata_017$`Valid Day Count`,na.rm = TRUE,trim=0.1)# na.rm ignores NA values
```





# VARIABILITY METRICS

### Variance
  - The sum of squared deviations from the mean divided by n - 1 where n is the number of data values.
  
```{r}

var(USEPAdata_017$`Valid Day Count`,na.rm = TRUE)# na.rm ignores NA values
```
  

### Standard deviation
  - The square root of the variance.
  
```{r}
sd(USEPAdata_017$`Valid Day Count`,na.rm = TRUE)# na.rm ignores NA values
```



### Mean absolute deviation
  - The mean of the absolute value of the deviations from the mean.
  
```{r}
MeanAD(USEPAdata_017$`Valid Day Count`,na.rm = TRUE)# na.rm ignores NA values
```
 
  
  
### Median absolute deviation from the median
  - The median of the absolute value of the deviations from the median.
  
```{r}
mad(USEPAdata_017$`Valid Day Count`,na.rm = TRUE)
```
  

### Range
  - The difference between the largest and the smallest value in a data set.
  
```{r}
range(USEPAdata_017$`Valid Day Count`,na.rm = TRUE)# na.rm ignores NA values

# diff gives difference between maximum and minimum values of the range

diff(range(USEPAdata_017$`Valid Day Count`,na.rm = TRUE))# na.rm ignores NA values
```
  

### Interquartile range
  - The difference between the 75th percentile and the 25th percentile.
  
  
```{r}
IQR(USEPAdata_017$`Valid Day Count`,na.rm = TRUE)
```
  
  
# EXPLORING THE DISTRIBUTION

### Boxplot
  - A plot introduced by Tukey as a quick way to visualize the distribution of data.
  
```{r}
boxplot(crim~rad,data=Boston,log="y",las=1,varwidth=TRUE,
        xlab="rad",
main ="Box Plot: Variation of boston crime values over discrete rad values")
```
  

### Frequency table
  - A tally of the count of numeric data values that fall into a set of intervals (bins).
  
```{r, message=FALSE,warning=FALSE}

  freq(Cars93$Manufacturer)
```


### Histogram
  - A plot of the frequency table with the bins on the x-axis and the count (or proportion) on the y-axis.
  
```{r}
hist(Cars93$Horsepower,
main="histogram")
```

### Density plot
  - A smoothed version of the histogram, often based on a kernel density estimate.
  
```{r}
# Create index16, pointing to 16-week chicks
index16<-which(ChickWeight$Time==16)

# Get the 16-week chick weights
weights <- (ChickWeight$weight[ChickWeight$Time==16])

# Plot the normalized histogram
truehist(weights)

# Add the density curve to the histogram
lines(density(weights))

```
  

#  CORRELATION

### Correlation coefficient
  - A metric that measures the extent to which numeric variables are associated with one another (ranges from -1 to +1).

### Correlation matrix
  - A table where the variables are shown on both rows and columns, and the cell values are the correlations between the variables.


```{r}

cor(mtcars)# Provides correlation values
corrplot(cor(mtcars),method="ellipse")
```


### Scatterplot
  - A plot in which the x-axis is the value of one variable, and the y-axis the value of another.
  
```{r}
plot(medv~rm,data=Boston)
title("Standard scatterplot")
```



**Referred from oreilly.com**
[Link](https://www.oreilly.com/library/view/practical-statistics-for/9781491952955/ch01.html)





