---
title: "Data Import R cheatsheet"
author: "Shanmukh"
date: "April 10, 2019"
output:
  html_document:
    toc: TRUE
    toc_depth: 3
    df_print: paged
---

reference: 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read Tabular Data

Install readr

```{r, echo=FALSE}
library(readr)
```


- read_* functions shares the common arguments:

syntax:

read_*(file, col_names = TRUE, col_types = NULL, locale = default_locale(), na = c("", "NA"),
quoted_na = TRUE, comment = "", trim_ws = TRUE, skip = 0, n_max = Inf, guess_max = min(1000,
n_max), progress = interactive())


### read_csv

- loads a comma separated file.

```{r, message=FALSE}
billboard <- read_csv("https://raw.githubusercontent.com/shapinna/Codeporfolio/master/Data/billboard.csv", col_names=TRUE)

head(billboard)[,c(1:5)]

```

### read_delim

- loads file with any delimiter
- syntax: read_delim("file.txt", delim = "|")

```{r,message=FALSE}


billb <- read_delim("https://raw.githubusercontent.com/shapinna/Codeporfolio/master/Data/billboard.csv", col_names = TRUE,delim = ",")

head(billb)[,c(1:5)]
```


### read.table

- read.table() loads text file.

```{r}

students <- read.table("http://users.stat.ufl.edu/~aa/smss/data/Students.dat")

head(students)[,c(1:9)]

```


### Other arguments

* Skip lines
    - Skips lines while loading
    - syntax: read_csv(f, skip = 1)
    
* No header
    - Loads file without headers
    - syntax: read_csv(f, col_names = FALSE)

* Provide header
    - Loads file with given headers
    - syntax: read_csv(f, col_names = c("x", "y", "z"))
    



# Read Non-Tabular Data

### Read a file into a single string

- read_file(file, locale = default_locale())


```{r}
f <- read_file("https://raw.githubusercontent.com/shapinna/Codeporfolio/master/Data/scraping.txt")

f

```

### Read each line into its own string

- read_lines(file, skip = 0, n_max = -1L, na = character(),locale = default_locale(), progress = interactive())


```{r}
text <- read_lines("https://raw.githubusercontent.com/shapinna/Codeporfolio/master/Data/scraping.txt")
text[1:3] # view first three lines of the file
```




*Referred from Rstudio cheatsheet*
[Link](https://github.com/rstudio/cheatsheets/raw/master/data-import.pdf)





