---
title: "Introduction and Intermediate R cheatsheet"
author: "Shanmukh"
date: "April 10, 2019"
output:
  html_document:
    toc: TRUE
    toc_depth: 4
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Introduction to R

#### Arithmetic with R

- Addition: +
- Subtraction: -
- Multiplication: *
- Division: /
- Exponentiation: ^
- Modulo: %%

```{r}
# An addition
2 + 2 

# A subtraction
2 - 2 

# A multiplication
3 * 2

 # A division
(3 + 3) / 2 

# Exponentiation
2^6

# Modulo
28 %% 5
```


#### Variable assignment


```{r}
# Assign 5 to variable X

X <- 5

```

#### Data types in R

```{r}
class(30)  # Numeric
class(30.5) # Numeric
class(TRUE)# Logical
class("Apple") # character


```


#### vector

-Vectors (one dimensional array): can hold numeric, character or logical values. The elements in a vector all have the same data type.

```{r}
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector
total_vector <- A_vector + B_vector


# Adding names to vector

week_vector <- c("WK1", "Wk2", "wk3")

names(total_vector) <- week_vector

# Print out total_vector
total_vector

```

#### Matrix

- a matrix is a collection of elements of the same data type (numeric, character, or logical) arranged into a fixed number of rows and columns

- a matrix is called two-dimensional
```{r}
# Construct a matrix with vector and provide row and column labels

ta <- c(1,2,3,4)
tb <- c(5,6,7,8,9)
tc <- c(ta,tb)

tm <- matrix(tc,byrow = TRUE, nrow=3)

tm  # display matrix
```

##### adding row and column labels

```{r}
tcln <- c("c1","c2","c3")
trn <- c("r1","r2","r3")
colnames(tm) <- tcln  # column label
rownames(tm) <- trn   # Row label

tm
```



##### adding totals to matrix rows and columns

```{r}
rtot <- rowSums(tm)
tmr <- cbind(tm,rtot)
tmr # with row total

ctot <- colSums(tm)
tmc <- rbind(tm,ctot)
tmc # with column total

```

#### Factor

```{r}
#nominal categorical variable

animals_vector <- c("Elephant", "Giraffe", "Donkey", "Horse")
factor_animals_vector <- factor(animals_vector)
factor_animals_vector

# ordinal categorical variable
# Temperature
temperature_vector <- c("High", "Low", "High","Low", "Medium")
factor_temperature_vector <- factor(temperature_vector, order = TRUE, levels = c("Low", "Medium", "High"))
factor_temperature_vector
```


```{r}
# Create speed_vector
speed_vector <- c("medium", "slow", "slow", "medium", "fast")

# Convert speed_vector to ordered factor vector
factor_speed_vector <-factor(speed_vector,
                             ordered= TRUE,
                             levels = c("slow","medium","fast"))

factor_speed_vector
```


#### List


```{r}
ex_list<- list(
  a=c(1,2,3,4),
  b= TRUE,
  c = "Hello"
)

ex_list[1] # displays 1st value

ex_list$a # display index with label a

ex_list[[1]]  # dsiplay only elements of 1st indext without label name
```




#### Data Frame

- Data frames (two-dimensional objects): can hold numeric, character or logical values. Within a column all elements have the same data type, but different columns can be of different data type.

```{r}
tdf <- data.frame(tc,cl2= 21:29,cl3=3)
head(tdf) # top 5 rows
tail(tdf) # Last 5 rows
str(tdf)  # structure of data frame


```





## Intermediate R



#### Relational operators

- ==  Equal comparison
- !=  Not equal comparison
- >   Greater than comparison
- <   Less than comparison
- >=
- <=

```{r}
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Popular days
linkedin>15

# Quiet days
linkedin<=5

# LinkedIn more popular than Facebook
linkedin>facebook

```

#### Logical operators

- & And operator
- Or operator
-! Not operator

```{r}

# or operator

linkedin<=5 | linkedin>15

# & operator

linkedin>=10 & linkedin<15

```

#### if and else if and else statement

syntax: 
if (condition1) {
  expr1
} else if (condition2) {
  expr2
} else if (condition3) {
  expr3
} else {
  expr4
}

```{r}

# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else if (medium == "Facebook"){
  # Add code to print correct string when condition is TRUE
  print("Showing Facebook information")
} else {
  print("Unknown medium")
}

# Control structure for num_views
if (num_views > 15) {
  print("You're popular!")
} else if (num_views <= 15 & num_views > 10) {
  # Add code to print correct string when condition is TRUE
  print("Your number of views is average")
} else {
  print("Try to be more visible!")
}


```

#### while loop

syntax:
while (condition) {
  expr
}

```{r}

i <- 1

# Code the while loop
while (i <= 10) {
  print(3*i)
  if (3*i%%8==0 ) {
    break
  }
  i <- i + 1
}


```

#### For Loop with break

```{r}
# Pre-defined variables
rquote <- "r's internals are irrefutably intriguing"
chars <- strsplit(rquote, split = "")[[1]]

# Initialize rcount
rcount <- 0

# Finish the for loop
for (char in chars) {
  if(char=="r"){
    rcount<- rcount+1
  }
  else if (char=="a"){
  break
}

# Print out rcount
print(rcount)
}
```


#### Functions
input -- process -- output


Syntax for creating functions

my_func <- function(arg1,arg2){
  body
}

```{r}
# Create a function pow_two()
pow_two <- function(a){
  y<-a*a
  return(y)
}


# Use the function
pow_two(12)

# Create a function sum_abs()
sum_abs<- function(a,b){
  y<-abs(a)+abs(b)
  return(y)
}


# Use the function
sum_abs(-2,3)
```


#### loading R package

- install.packages(), which as you can expect, installs a given package.
- library() which loads packages, i.e. attaches them to the search list on your R workspace.

```{r}
library(MASS)
# Check out the currently attached packages again
search()
```

#### strsplit
syntax: strsplit(x, split, fixed = FALSE, perl = FALSE, useBytes = FALSE)
```{r}
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split_math <- strsplit(pioneers, split = ":")
split_math
```

#### lapply

lapply takes a vector or list X, and applies the function FUN to each of its members. If FUN requires additional arguments, you pass them after you've specified X and FUN (...). The output of lapply() is a list, the same length as X, where each element is the result of applying FUN on the corresponding element of X.
```{r}
split_low <- lapply(split_math,tolower)

split_low

pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Generic select function
select_el <- function(x, index) {
  x[index]
}

# Use lapply() twice on split_low: names and years
names<- lapply(split_low,select_el,1)
years<- lapply(split_low,select_el,2)

names
years

```

#### sapply

simplified version of sapply it will try to generate to array if possible otherwise leave as list

```{r}

l1 <- list(c(3,7,9,6,-1))
l2 <- list(c(6,9,12,13,5))

temp <-c(l1,l2)


# Use lapply() to find each day's minimum temperature
lapply(temp,min)

# Use sapply() to find each day's minimum temperature
sapply(temp,min)

# Use lapply() to find each day's maximum temperature
lapply(temp,max)

# Use sapply() to find each day's maximum temperature
sapply(temp,max)


```

##### vapply
same as sapply but output can be specified
Vapply throws error if the expected output is not returned

#### R functions

* abs -- calculates absolute value of numerical values

* round -- rounds the integers.

* sum -- sums up all the given values

* c -- Combines two or more specific values

* mean -- calculate mean of the values
```{r}
v1 <- c(1.1,-5.1,2.5,-6.4)
v2 <- c(3.1,2.3,-6.1,-9.9)

sum(round(abs(v1)))
sum(round(abs(v2)))

mean(c(sum(round(abs(v1))),sum(round(abs(v2)))))

```

* seq -- sequence of numbers generated by r by parameters given in the argument

* rep -- repeats the values

* sort -- sort the values

```{r}
li <- list(
  log = TRUE,
  ch = "hello",
  int_vec = sort(rep(seq(8,2, by=-2),times=2))
)

str(li)
```

* is.*() -- check whether given variable of same data type and returns result in  logical value

* as.*()-- converts the data type

* unlist() -- unlist the list function to a vector with lables

```{r}
is.list(li)

is.list(c(1,2,3))

li2 <- as.list(c(1,2,3))

is.list(li2)

unlist(li)

```
* append-- allows to add elements to vector or a list 
* rev -- reverses elements in a data structure

```{r}
str(append(li,rev(li)))

```


#### Times and dates

```{r}
today<-Sys.Date()  # todays date
now <- Sys.time() # current time

str(today)
str(now)

my_date <- as.Date("2010-15-05",format="%Y-%d-%m")  # Default format %Y-%m-%d
str(my_date)

#Arithmetics

New_date2 <- as.Date("2005-06-06")

my_date-New_date2 # time difference from two dates

unclass(today) # today value stored in R it is no of days from 1 jan 1970

unclass(now) # current time value stored in R


```

* advanced date packages are lubridate, zoo, xts

* as.date formats:
  - %Y: 4-digit year (1982)
  - %y: 2-digit year (82)
  - %m: 2-digit month (01)
  - %d: 2-digit day of the month (13)
  - %A: weekday (Wednesday)
  - %a: abbreviated weekday (Wed)
  - %B: month (January)
  - %b: abbreviated month (Jan)

* as.POSIXct formats:
  - %H: hours as a decimal number (00-23)
  - %I: hours as a decimal number (01-12)
  - %M: minutes as a decimal number
  - %S: seconds as a decimal number
  - %T: shorthand notation for the typical format %H:%M:%S
  - %p: AM/PM indicator

```{r}
x <- list(2006, 2004, 2020, 2002)
for(i in 1:length(x)) {
    if(x[[i]] %% 4 != 0) next
  print(paste(x[[i]], "is a leap year"))
}
```


