---
title: "Regression Analysis and Clustering"
author: "Shanmukh"
date: "April 10, 2019"
output:
  html_document:
    toc: TRUE
    toc_depth: 4
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear regression

```{r,echo=FALSE,warning=FALSE,message=FALSE}

library(car) # test for heteroscedasticity
library(purrr) # for cluster analysis
```


#### Linear regression equation

```{r}
mdl <- lm(mpg ~ hp, data = mtcars)# Linear regression equation

mdl # Intercept and beta co-efficient values

```



#### summary of linear regression equation

```{r}
summary(mdl)
```



#### adding interaction term

```{r}
# Linear regression model with interaction term

mdl1 <- lm(mpg ~ hp+am, data = mtcars)

summary(mdl1)

```

#### Verify the linear relationship between through scatter plot with trend line

```{r}

plot(mtcars$mpg~mtcars$hp,
     col="darkblue",pch=18,
     ylab="mpg",
     xlab="hp")
abline(mdl,lwd=2,col="red")
```

#### Verification of outliers:

```{r}
# cooks distance

cd_mdl1<-cooks.distance(mdl1)

sum(cd_mdl1 > 4 / length(cd_mdl1))

plot(mdl1,which = 4)  #Cooks distance

```

#### QQ Plot:

```{r}
plot(mdl1,which = 2)

#Alternate option1
qqPlot(mdl1)

# alternate option2
qqnorm(resid(mdl1), main = "Normal Q-Q Plot, mdl1", col = "darkgrey")
qqline(resid(mdl1), col = "dodgerblue", lwd = 2)
```




#### Test for heteroscedasticity

-- Verify the variability in the residuals is same for both large and small values for estimated Y

##### Plot of residuals vs predicted values

```{r}

# Plot for residuals Vs fitted values

plot(mdl1$residuals~mdl1$fitted.values,main="Residuals vs Predicted Values")
abline(h=0,col="red",lty=2,lwd=2) 




plot(mdl1,which=4)


residuals(mdl1)
fitted(mdl1)

```

#####	Statistical test for heteroscedasticity

-- if P value is > alpha we can assume there is no heteroscedasticity

```{r}
ncvTest(mdl1)
```

#### verify Residuals to be normally distributed

```{r}
hist(mdl1$residuals,col="grey75",main="Distribution of Residuals",prob=TRUE)
curve(dnorm(x,mean(mdl1$residuals),sd(mdl1$residuals)),add=TRUE,col="red")
```

#### shapiro.test

```{r}

shapiro.test(resid(mdl1))
```

```{r}
str(ChickWeight)

# Get the 16-week chick weights
weights <- (ChickWeight$weight[ChickWeight$Time==16])

library(MASS)
# Plot the normalized histogram
truehist(weights)
# Add the density curve to the histogram
lines(density(weights))

library(car)
qqPlot(weights)

qqPlot(Boston$tax)

```



## Cluster analysis

**mtcars dataset**
```{r}
str(mtcars)
```


#### Elbow point

```{r}
wss <- function(k) {
  kmeans(mtcars, k, nstart = 25 )$tot.withinss
}

# Compute and plot wss for k = 1 to k = 15
k.values <- 1:31

# extract wss for 2-15 clusters
wss_values <- map_dbl(k.values, wss)

plot(k.values, wss_values,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")


```



```{r}


tot_withinss1 <- map_dbl(1:31,function(k){
  model <- kmeans(mtcars,centers=k)
  model$tot.withinss
})
elbow_df <-data.frame(
  k=1:31,
  tot_withinss1=tot_withinss1
)

print(elbow_df)


```

