---
title: "Regular expressions r cheatsheet"
author: "Shanmukh"
date: "April 10, 2019"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    df_print: paged

---
Install library
```{r echo=FALSE, warning=FALSE, message = FALSE}
library(stringr)# for string manipulation
library(htmlwidgets) # for html widgets
library(tidyverse) # for data wrangling
library(readr) # for loading the data
```

# Regular expressions

* grep -- returns index values of the elements it matched to the regrular expression pattern 

* grep1 -- returns a logical values if the provided regrular expression pattern is found

* syntax  grepl(pattern=<regex>,x= <string>)


```{r}
animals <- c("cat","moose","impala","ant","kiwi")

grepl("a",animals) # checks for contains a

grepl("^a",animals) #^a checks for starting with a

grepl("a$",animals) # a$ checks for ending with a

grep("a",animals) # checks for contains a

which(grepl("a",animals))

```

* sub-- replaces first match value with replacement string

* syntax  sub(pattern= <regex>, replacement=<string>,x=<string>)

* gsub-- replaces all matched values with replacement string

```{r}
sub("a","b",animals) # replaces first a

gsub("a","b",animals) # replaces all a in the string

gsub("a|i","_",animals) # replaces all a or i with _ in the string

```
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```

**Import text file**
```{r}
text <- read_lines("https://raw.githubusercontent.com/shapinna/Codeporfolio/master/Data/scraping.txt")
text[1:3] # view first three lines of the file
```

**extract string - text**
```{r}
str_extract_all(text, "[0-9]+")[[1]] # list with numbers from the text file
```

**build regex**
`[a-zA-Z]+\s[0-9][0-9],\s[0-9]{4}`

```{r}
dates <- regex("
[a-zA-Z]+ # month
\\s       # Space
[0-9]     # 1 digit
[0-9]     # 1 digit
,         # comma
\\s       # Space
[0-9]{4}  # 4 digits
", comments = TRUE)  # comments are valid with in the regex function
```
**Match regex - print all**
```{r}
str_match(text, dates) # regular expression format stored in the dates is compared with each line of the text file and NA is returned if matching records does not exists.
```
**Print only dates**
```{r}
# for loop to print only matching record exists
results = str_match(text, dates)
for (i in results){
  if (!is.na(i)){
    print(i)
  }
}
```