---
title: "Visualization R cheatsheet"
author: "Shanmukh"
date: "April 10, 2019"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Base R Visualizations

# Data visualization in R
```{r}
str(mtcars)
ss_mt<- mtcars[,c(1,3,5)]
plot(ss_mt)
```

#### Scatter plot
Syntax:  Plot(yvariable~Xvariable, data= data)

```{r echo=FALSE, warning=FALSE, message = FALSE} 
library(MASS)

plot(medv~rm,data=Boston)

sunflowerplot(tax~rad,data=Boston)

boxplot(crim~rad,data=Boston,log="y",las=1,varwidth=TRUE,
        xlab="rad",
main ="Variation of boston crime values over discrete rad values")

mosaicplot(cyl~gear,data=mtcars, main="Mosaic plot")



```

```{r}
# Set up a side-by-side plot array
par(mfrow=c(1,2))

# Create the standard scatterplot
plot(rad~zn,data=Boston)

# Add the title
title("Standard scatterplot")

# Create the sunflowerplot
sunflowerplot(rad~zn,data=Boston)

# Add the title
title("Sunflower plot")
```
#### Bagplots : two dimensional box plots

```{r}
library(aplpack)
par(mfrow=c(1,2))
boxplot(Cars93$Min.Price,Cars93$Max.Price)
bagplot(Cars93$Min.Price,Cars93$Max.Price,cex=1.2)
abline(a=0,b=1,lty=2)

```




#### correlation plots
syntax: corrplot(cor(dataset),method="ellipse")
```{r}
library(corrplot)
cor(mtcars)

corrplot(cor(mtcars),method="ellipse")


```


```{r}

# Set up a 2-by-2 plot array
par(mfrow=c(2,2))

index<- which(ChickWeight$Time==16)

# Plot the Animals2 brain weight data as points
plot(ChickWeight$weight[index], type="p")

# Add the title
title("points")

# Plot the brain weights with lines
plot(ChickWeight$weight[index],type="l")

# Add the title
title("lines")

# Plot the brain weights as lines overlaid with points
plot(ChickWeight$weight[index],type="o")

# Add the title
title("overlaid")

# Plot the brain weights as steps
plot(ChickWeight$weight[index],type="s")

# Add the title
title("steps")
```

```{r}
# Create MPG.city vs. Horsepower plot with solid squares
plot(MPG.city~Horsepower,data=Cars93,pch=15)

#plot(ChickWeight$weight[index], type="p")

# Create index3, pointing to 3-cylinder cars
index3=which(Cars93$Cylinders==3)

# Add text giving names of cars next to data points
text(x =Cars93$Horsepower[index3] , 
     y =Cars93$MPG.city[index3] ,
     labels = Cars93$Make[index3], adj = 0)
```
```{r}
# Plot MPG.city vs. Horsepower as open circles
plot(MPG.city~Horsepower,data=Cars93)

# Create index3, pointing to 3-cylinder cars
index3 = which(Cars93$Cylinders==3)

# Highlight 3-cylinder cars as solid circles
points(Cars93$Horsepower[index3], Cars93$MPG.city[index3],pch=16)

# Add car names, offset from points, with larger bold text
text(Cars93$Horsepower[index3], Cars93$MPG.city[index3],label=Cars93$Make[index3],adj=-0.2,cex=1.2,font=2)
```
```{r}
# Plot Gas vs. Temp as solid triangles
plot(Gas~Temp,data=whiteside,pch=17)

# Create indexB, pointing to "Before" data
indexB<- which(whiteside$Insul=="Before")

# Create indexA, pointing to "After" data
indexA<- which(whiteside$Insul=="After")

# Add "Before" text in blue, rotated 30 degrees, 80% size
text(x = whiteside$Temp[indexB], y = whiteside$Gas[indexB],
     labels = "Before", col = "blue", srt = 30, cex = 0.8)

# Add "After" text in red, rotated -20 degrees, 80% size
text(x = whiteside$Temp[indexA], y = whiteside$Gas[indexA],
     labels = "After", col = "red", srt = -20, cex = 0.8)
```
```{r}
# Set up and label empty plot of Gas vs. Temp
plot(whiteside$Temp,whiteside$Gas,
     type = "n", xlab ="Outside temperature",
     ylab = "Heating gas consumption")

# Create indexB, pointing to "Before" data
indexB <- which(whiteside$Insul=="Before")

# Create indexA, pointing to "After" data
indexA <- which(whiteside$Insul=="After")

# Add "Before" data as solid triangles
points(whiteside$Temp[indexB],whiteside$Gas[indexB],pch=17)

# Add "After" data as open circles
points(whiteside$Temp[indexA],whiteside$Gas[indexA],pch=1)

# Add legend that identifies points as "Before" and "After"
legend("topright", pch =c(17,1), 
       legend = c("Before", "After"))
```

```{r}
library(MASS)

# Create a boxplot of sugars by shelf value, without axes
boxplot(sugars~shelf,data=UScereal,axes=FALSE)

# Add a default y-axis to the left of the boxplot
axis(side =2,at=UScereal$sugars, labels=as.character(UScereal$sugars))

# Add an x-axis below the plot, labelled 1, 2, and 3
axis(side = 1, at = 1:3, labels=c(1,2,3))

# Add a second x-axis above the plot
  axis(side = 3, at = 1:3,
       labels = c("floor","middle","top"))

#is.factor(UScereal$shelf)

#str(UScereal)
```

```{r}
# Create a scatterplot of MPG.city vs. Horsepower
plot(Cars93$Horsepower,Cars93$MPG.city)

# Call supsmu() to generate a smooth trend curve, with default bass
trend1<-supsmu( Cars93$Horsepower,Cars93$MPG.city,  bass=0)

# Add this trend curve to the plot
lines(trend1)

# Call supsmu() for a second trend curve, with bass = 10
trend2 <- supsmu( Cars93$Horsepower,Cars93$MPG.city,  bass=10)

# Add this trend curve as a heavy, dotted line
lines(trend2,lty=3,lwd=2)
```

```{r}

# Create mfr_table of manufacturer frequencies
mfr_table <- table(Cars93$Manufacturer)

# Create the default wordcloud from this table
wordcloud(words = names(mfr_table), 
          freq = as.numeric(mfr_table), 
          scale = c(2,0.25))

# Change the minimum word frequency
wordcloud(words = names(mfr_table), 
          freq = as.numeric(mfr_table), 
          scale = c(2,0.25), 
          min.freq = 3)

# Create model_table of model frequencies
model_table <- table(Cars93$Model)

# Create the wordcloud of all model names with smaller scaling
wordcloud(words = names(model_table), 
          freq = as.numeric(model_table), 
          scale = c(0.75, 0.25), 
          min.freq = 1)

```
```{r}
# Create row1, row2, and layoutVector
row1 <- c(1,0,0)
row2 <- c(0,2,2)
layoutVector <- c(row1,row2,row2)

# Convert layoutVector into layoutMatrix
layoutMatrix <- matrix(layoutVector, byrow = TRUE, nrow = 3)

# Set up the plot array
layout(layoutMatrix)

# Plot scatterplot
plot(Boston$rad,Boston$zn)

# Plot sunflower plot
sunflowerplot(Boston$rad,Boston$zn)
```


```{r}
# Create a table of Cylinders frequencies
tbl <- table(Cars93$Cylinders)

# Generate a horizontal barplot of these frequencies
mids <- barplot(tbl, horiz = TRUE, col = "transparent",
                names.arg = "")

# Add names labels with text()
text(20, mids, names(tbl))

# Add count labels with text()
text(35, mids, as.numeric(tbl))
```

```{r}
# Call symbols() to create the default bubbleplot
symbols(Cars93$Horsepower,Cars93$MPG.city,
        circles = Cars93$Price)

# Repeat, with the inches argument specified
symbols(Cars93$Horsepower,Cars93$MPG.city,
        circles = Cars93$Price,
        inches = 0.1)
```

```{r}
# Call png() with the name of the file we want to create
png("bubbleplot.png")

# Re-create the plot from the last exercise
symbols(Cars93$Horsepower, Cars93$MPG.city,
        circles = sqrt(Cars93$Price),
        inches = 0.1)

# Save our file and return to our interactive session
dev.off()

# Verify that we have created the file
list.files(pattern = "png")
```

```{r}
# Iliinsky and Steele color name vector
IScolors <- c("red", "green", "yellow", "blue",
              "black", "white", "pink", "cyan",
              "gray", "orange", "brown", "purple")

# Create the data for the barplot
barWidths <- c(rep(2, 6), rep(1, 6))

# Recreate the horizontal barplot with colored bars
barplot(rev(barWidths), horiz = TRUE,
        col = rev(IScolors), axes = FALSE,
        names.arg = rev(IScolors), las = 1)

```
```{r}
# Iliinsky and Steele color name vector
IScolors <- c("red", "green", "yellow", "blue",
              "black", "white", "pink", "cyan",
              "gray", "orange", "brown", "purple")

# Create the `cylinderLevels` variable
cylinderLevels<- Cars93$Cylinders

# Create the colored bubbleplot
symbols(Cars93$Horsepower,Cars93$MPG.city, 
        circles = cylinderLevels, inches = 0.2, 
        bg = IScolors[cylinderLevels])
```

```{r}
# Create a table of Cylinders by Origin
tbl <- table(Cars93$Cylinders,Cars93$Origin)

# Create the default stacked barplot
barplot(tbl,names.arg=names(tbl))


# Enhance this plot with color
barplot(tbl,names.arg=names(tbl),col=IScolors)
```



# ggplot

Install **ggplot2** package

```{r,echo=FALSE,message=FALSE,warning=FALSE}

library(ggplot2)
mpg

```
ggplot syntax
ggployt(data= DATA)+
<GEOM_FUNCTION>(mapping= aes(MAPPING))
```{r}
ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,color=class))
```

```{r}
ggplot(mpg)+
  geom_point(mapping=aes(x=hwy,y=cyl))
```

```{r}
ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,size=class))

ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,alpha=class))

ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,shape=class))

ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy),color="blue")
```
# Facets

Facets is particular useful for categorical variable, it is used to split the plot in to subplots each display subset of data

Facet_wrap: first argument in facetwrap should be a formula which is created by ~ and a variable

Facet_grid: Facet grid should contain two ariable names separated by ~
```{r}
ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  facet_wrap(~class,nrow = 2)

ggplot(mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))+
  facet_grid(drv~cyl)

ggplot(mpg)+
  geom_point(mapping=aes(x=drv,y=cyl))


```
# Geom objects
Geom is a geometrical object that a plot uses to represent data

bar charts use bar geioms
line charts use line geoms box plots use box plot geom

```{r}
ggplot(mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))

ggplot(mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy))

ggplot(mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy,linetype=drv))

ggplot(mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy,group=drv))

ggplot(mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy,color=drv),
              show.legend = FALSE)

```
# multiple geoms in single graph
```{r}
ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  geom_smooth(mapping=aes(x=displ,y=hwy))

ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point()+
  geom_smooth()

# extend mapping in each geom
ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point(mapping = aes(color=class))+
  geom_smooth()

# filter data only for a geom
ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point(mapping = aes(color=class))+
  geom_smooth(data=filter(mpg,class=="subcompact"),se=FALSE)


ggplot(mpg, aes(x = displ, y = hwy)) +
 geom_point() +
 geom_point(data = filter(mpg, class=="2seater"), color = 'red')

```
# statistical transformations

```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut))

## bar chart with proportion
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,y=..prop..,group=1))

## summary stat

ggplot(data=diamonds)+
  stat_summary(mapping=aes(x=cut,y=depth),
               fun.ymin = min,
               fun.ymax = max,
               fun.y=median)

```
# position adjustments
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,color=cut))

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=cut))

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity))

# position fill works a proportion across groups
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="fill")

# position dodge places overlapping objects directly beside one another

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="dodge")

```
# coordinate system

Default coordinate system in ggplot is cartesian coordinate system where x and y positions act independently


-coordflip switches x and y axes
```{r}

ggplot(data=mpg,mapping=aes(x=class,y=hwy))+
  geom_boxplot()

ggplot(data=mpg,mapping=aes(x=class,y=hwy))+
  geom_boxplot()+
  coord_flip()


```

ggplot(data=nz,aes(long,lat,group=group))+
  geom_polygon(fill="white",color="black")+
  coord_quickmap()




```{r}
bar <- ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=cut),show.legend = FALSE,
           width=1)+
  theme(aspect.ratio = 1)+
  labs(x=NULL,y=NULL)

bar+coord_flip()
bar+coord_polar()
```
ggplot syntax
ggployt(data= DATA)+
<GEOM_FUNCTION>(mapping= aes(MAPPING),
stat=<STAT>,
position=<POSITION>)+
<COORDINATE_FUNCTION>+
<FACET_FUNCTION>


```{r}



# Create the basic plot (not displayed): basePlot
basePlot <- ggplot(Cars93, aes(x = Horsepower, y = MPG.city))

# Display the basic scatterplot
basePlot + 
  geom_point()

# Color the points by Cylinders value
basePlot + 
  geom_point(color= IScolors[Cars93$Cylinders])

# Make the point sizes also vary with Cylinders value
basePlot + 
  geom_point(color = IScolors[Cars93$Cylinders], 
             size = as.numeric(Cars93$Cylinders))
```