---
title: "Data visualization in R cheatsheet"
author: "Shanmukh"
date: "April 10, 2019"
output:
  html_document:
    toc: TRUE
    toc_depth: 3
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Base R Visualizations

```{r,warning=FALSE,echo=FALSE,message=FALSE}
library(MASS)
library(dplyr)
library(maps)
library(aplpack)
library(corrplot)
```

Dataset: mtcars

```{r echo=FALSE}
head(mtcars)
```

```{r}
ss_mt<- mtcars[,c(1,3,5)]
plot(ss_mt)
```



### Scatter plot
Syntax:  Plot(yvariable~Xvariable, data= data)

```{r } 

plot(medv~rm,data=Boston)
title("Standard scatterplot")

```
### Sunflower plot
Syntax:  sunflowerplot(yvariable~Xvariable, data= data)
```{r}
sunflowerplot(tax~rad,data=Boston)
title("Sunflower plot")
```
### Box Plot
syntax: boxplot(yvariable~Xvariable, data= data,log="y")
```{r}
boxplot(crim~rad,data=Boston,log="y",las=1,varwidth=TRUE,
        xlab="rad",
main ="Box Plot: Variation of boston crime values over discrete rad values")
```
### mosaicplot
syntax: mosaicplot(yvariable~Xvariable, data= data)
```{r}
mosaicplot(cyl~gear,data=mtcars, main="Mosaic plot")

```

### Histogram vs true hist

- hist() is part of base R and its default option yields a histogram based on the number of times a record falls into each of the bins on which the histogram is based.

- truehist() is from the MASS package and scales these counts to give an estimate of the probability density.

```{r}
par(mfrow=c(1,2))

hist(Cars93$Horsepower,
main="hist() plot")

truehist(Cars93$Horsepower,
main="truehist() plot")

```


### Box plot vs Bagplots:

- A single box plot gives a graphical representation of the range of variation in a numerical variable, based on five numbers:

The minimum and maximum values
The median (or "middle") value
Two intermediate values called the lower and upper quartiles

- The bag plot extends this representation to two numerical variables, showing their relationship, both within two-dimensional "bags" corresponding to the "box" in the standard boxplot, and indicating outlying points outside these limits.



```{r}
par(mfrow=c(1,2))  # Set up a side-by-side plot array
boxplot(Cars93$Min.Price,Cars93$Max.Price)
bagplot(Cars93$Min.Price,Cars93$Max.Price,cex=1.2)
abline(a=0,b=1,lty=2)

```

### Density plots

-  The density plot of this type of variable is often expected to conform approximately to the bell-shaped curve, otherwise known as the Gaussian distribution

```{r}
# Create index16, pointing to 16-week chicks
index16<-which(ChickWeight$Time==16)

# Get the 16-week chick weights
weights <- (ChickWeight$weight[ChickWeight$Time==16])

# Plot the normalized histogram
truehist(weights)

# Add the density curve to the histogram
lines(density(weights))
```

### correlation plots
syntax: corrplot(cor(dataset),method="ellipse")
```{r}

cor(mtcars)# Provides correlation values
corrplot(cor(mtcars),method="ellipse")

```


### type option in plots: 

* The specific type values used here are:

    - "p" for "points"
    - "l" for "lines"
    - "o" for "overlaid" (i.e., lines overlaid with points)
    - "s" for "steps"

```{r}


par(mfrow=c(2,2))# Set up a 2-by-2 plot array  
index<- which(ChickWeight$Time==16)


plot(ChickWeight$weight[index], type="p") # Plot the Animals2 brain weight data as points
title("points")

plot(ChickWeight$weight[index],type="l") # Plot the brain weights with lines
title("lines")


plot(ChickWeight$weight[index],type="o") # Plot the brain weights as lines overlaid with points
title("overlaid")

plot(ChickWeight$weight[index],type="s")# Plot the brain weights as steps
title("steps")
```

### points()

```{r}

par(mfrow=c(1,2))
# Create an empty plot using type = "n"
plot(mtcars$hp,mtcars$mpg,type="n",
xlab="Horsepower",ylab="Gas mileage")

# Add points with shapes determined by cylinder number
points(mtcars$hp,mtcars$mpg, pch = mtcars$cyl)

# Create a second empty plot
plot(mtcars$hp,mtcars$mpg,type="n",
xlab="Horsepower",ylab="Gas mileage")

# Add points with shapes as cylinder characters
points(mtcars$hp,mtcars$mpg, pch =as.character(mtcars$cyl))
```


### text()

- text() function is to add informative labels to a data plot. 

- The text() function takes three arguments:
    - x, which specifies the value for the x variable,
    - y, which specifies the value for the y variable, and
    - label, which specifies the label for the x-y value pair.


```{r}
# Plot MPG.city vs. Horsepower as open circles
plot(MPG.city~Horsepower,data=Cars93)

# Create index3, pointing to 3-cylinder cars
index3 = which(Cars93$Cylinders==3)

# Highlight 3-cylinder cars as solid circles
points(Cars93$Horsepower[index3], Cars93$MPG.city[index3],pch=16)

# Add car names, offset from points, with larger bold text
text(Cars93$Horsepower[index3], Cars93$MPG.city[index3],label=Cars93$Make[index3],adj=-0.2,cex=1.2,font=2)
```

### legend()

- legend() function to add explanatory text to a plot.

```{r}
# Set up and label empty plot of Gas vs. Temp
plot(whiteside$Temp,whiteside$Gas,
     type = "n", xlab ="Outside temperature",
     ylab = "Heating gas consumption")

# Create indexB, pointing to "Before" data
indexB <- which(whiteside$Insul=="Before")

# Create indexA, pointing to "After" data
indexA <- which(whiteside$Insul=="After")

# Add "Before" data as solid triangles
points(whiteside$Temp[indexB],whiteside$Gas[indexB],pch=17)

# Add "After" data as open circles
points(whiteside$Temp[indexA],whiteside$Gas[indexA],pch=1)

# Add legend that identifies points as "Before" and "After"
legend("topright", pch =c(17,1), 
       legend = c("Before", "After"))
```

### axis():

- suppress the default axes when we create the plot by specifying axes = FALSE
- call the low-level graphics function axis() to create the axes.

```{r}
library(MASS)

# Create a boxplot of sugars by shelf value, without axes
boxplot(sugars~shelf,data=UScereal,axes=FALSE)

# Add a default y-axis to the left of the boxplot
axis(side =2,at=UScereal$sugars, labels=as.character(UScereal$sugars))

# Add an x-axis below the plot, labelled 1, 2, and 3
axis(side = 1, at = 1:3, labels=c(1,2,3))

# Add a second x-axis above the plot
  axis(side = 3, at = 1:3,
       labels = c("floor","middle","top"))

```

### supsmu():

- To add smooth trend curves

```{r}
# Create a scatterplot of MPG.city vs. Horsepower
plot(Cars93$Horsepower,Cars93$MPG.city)

# Call supsmu() to generate a smooth trend curve, with default bass
trend1<-supsmu( Cars93$Horsepower,Cars93$MPG.city,  bass=0)

# Add this trend curve to the plot
lines(trend1)

# Call supsmu() for a second trend curve, with bass = 10
trend2 <- supsmu( Cars93$Horsepower,Cars93$MPG.city,  bass=10)

# Add this trend curve as a heavy, dotted line
lines(trend2,lty=3,lwd=2)
```

### Layout matrix:


```{r}
# Create row1, row2, and layoutVector
row1 <- c(1,0,0)
row2 <- c(0,2,2)
layoutVector <- c(row1,row2,row2)

# Convert layoutVector into layoutMatrix
layoutMatrix <- matrix(layoutVector, byrow = TRUE, nrow = 3)

# Set up the plot array
layout(layoutMatrix)

# Plot scatterplot
plot(Boston$rad,Boston$zn)

# Plot sunflower plot
sunflowerplot(Boston$rad,Boston$zn)
```

#### Some plot functions also return useful information

```{r}
# Create a table of Cylinders frequencies
tbl <- table(Cars93$Cylinders)

# Generate a horizontal barplot of these frequencies
mids <- barplot(tbl, horiz = TRUE, col = "transparent",
                names.arg = "")

# Add names labels with text()
text(20, mids, names(tbl))

# Add count labels with text()
text(35, mids, as.numeric(tbl))
```

###  symbols()

- symbols() function allows us to extend scatterplots to show the influence of other variables.


```{r}
par(mfrow=c(1,2))
# Call symbols() to create the default bubbleplot
symbols(Cars93$Horsepower,Cars93$MPG.city,
        circles = Cars93$Price)

# Repeat, with the inches argument specified
symbols(Cars93$Horsepower,Cars93$MPG.city,
        circles = Cars93$Price,
        inches = 0.1)
```

### Saving plot results

```{r}
# Call png() with the name of the file we want to create
png("bubbleplot.png")

# Re-create the plot from the last exercise
symbols(Cars93$Horsepower, Cars93$MPG.city,
        circles = sqrt(Cars93$Price),
        inches = 0.1)

# Save our file and return to our interactive session
dev.off()

# Verify that we have created the file
list.files(pattern = "png")
```

#### Recommended colors

```{r}
# Iliinsky and Steele color name vector
IScolors <- c("red", "green", "yellow", "blue",
              "black", "white", "pink", "cyan",
              "gray", "orange", "brown", "purple")

# Create the data for the barplot
barWidths <- c(rep(2, 6), rep(1, 6))

# Recreate the horizontal barplot with colored bars
barplot(rev(barWidths), horiz = TRUE,
        col = rev(IScolors), axes = FALSE,
        names.arg = rev(IScolors), las = 1)

```

### bubbleplot

```{r}
# Iliinsky and Steele color name vector
IScolors <- c("red", "green", "yellow", "blue",
              "black", "white", "pink", "cyan",
              "gray", "orange", "brown", "purple")

# Create the `cylinderLevels` variable
cylinderLevels<- Cars93$Cylinders

# Create the colored bubbleplot
symbols(Cars93$Horsepower,Cars93$MPG.city, 
        circles = cylinderLevels, inches = 0.2, 
        bg = IScolors[cylinderLevels])
```

### stacked barplots


```{r}
par(mfrow=c(1,2))
# Create a table of Cylinders by Origin
tbl <- table(Cars93$Cylinders,Cars93$Origin)

# Create the default stacked barplot
barplot(tbl,names.arg=names(tbl))


# Enhance this plot with color
barplot(tbl,names.arg=names(tbl),col=IScolors)
```

# ggplot

Install **ggplot2** package

ggplot syntax
ggployt(data= DATA)+
<GEOM_FUNCTION>(mapping= aes(MAPPING),
stat=<STAT>,
position=<POSITION>)+
<COORDINATE_FUNCTION>+
<FACET_FUNCTION>

```{r,echo=FALSE,message=FALSE,warning=FALSE}

library(ggplot2)
str(mpg)

table(mpg$class)
```
### Creating a ggplot

- ggplot() creates a coordinate system that you can add layers to ( This creates an empty graph)
- geom_point() adds a layer of points to your plot, which creates a scatterplot

ggplot(data= DATA)+
<GEOM_FUNCTION>(mapping= aes(MAPPING))

```{r}
ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,color=class))
```



```{r}
ggplot(mpg)+
  geom_point(mapping=aes(x=hwy,y=cyl))
```

### Aesthetic mappings

* An aesthetic is a visual property of the objects in your plot, aesthetic includes
    - Size of the points
    - shape of the points
    - Color of the points


```{r, warning=FALSE}
ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,size=class)) # display points as size of class

ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,alpha=class)) # display points as alpha of class

ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,shape=class)) # display points as shape of class

ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy),color="blue") # display points color as blue
```

```{r}
# Create the basic plot (not displayed): basePlot
basePlot <- ggplot(Cars93, aes(x = Horsepower, y = MPG.city))

# Display the basic scatterplot
basePlot + 
  geom_point()

# Color the points by Cylinders value
basePlot + 
  geom_point(color= IScolors[Cars93$Cylinders])

# Make the point sizes also vary with Cylinders value
basePlot + 
  geom_point(color = IScolors[Cars93$Cylinders], 
             size = as.numeric(Cars93$Cylinders))
```




### Facets

Facets is particular useful for categorical variable, it is used to split the plot in to subplots each display subset of data

Facet_wrap: first argument in facetwrap should be a formula which is created by ~ and a variable

Facet_grid: Facet grid should contain two ariable names separated by ~
```{r}
ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  facet_wrap(~class,nrow = 2)

ggplot(mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))+
  facet_grid(drv~cyl)

ggplot(mpg)+
  geom_point(mapping=aes(x=drv,y=cyl))


```

### Geom objects

- Geom is a geometrical object that a plot uses to represent data

  - bar charts use bar geoms
  - line charts use line geoms
  - box plots use box plot geom

```{r}
ggplot(mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))
  
# a smooth line fitted to the data
ggplot(mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy))

# geom_smooth() will draw a different line, with a different linetype, for each unique value of the variable that you map to linetype

ggplot(mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy,linetype=drv))


ggplot(mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy,group=drv))

ggplot(mpg)+
  geom_smooth(mapping=aes(x=displ,y=hwy,color=drv),
              show.legend = FALSE)

```

### multiple geoms in single graph

```{r message=FALSE}
ggplot(mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  geom_smooth(mapping=aes(x=displ,y=hwy))

ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point()+
  geom_smooth()

# extend mapping in each geom
ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
  geom_point(mapping = aes(color=class))+
  geom_smooth()

# filter data only for a geom
ggplot(data=mpg,mapping=aes(x=displ,y=hwy))+
 geom_point(mapping = aes(color=class))+
 geom_smooth(data=filter(mpg,class=="subcompact"),se=FALSE)


ggplot(mpg, aes(x = displ, y = hwy)) +
 geom_point() +
 geom_point(data = filter(mpg, class=="2seater"), color = 'red')

```

### statistical transformations


```{r}
# Bar chart
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut))

## bar chart with proportion
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,y=..prop..,group=1))

## summary stat

ggplot(data=diamonds)+
  stat_summary(mapping=aes(x=cut,y=depth),
               fun.ymin = min,
               fun.ymax = max,
               fun.y=median)

```

### position adjustments
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,color=cut))

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=cut))

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity))

# position fill works a proportion across groups
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="fill")

# position dodge places overlapping objects directly beside one another

ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="dodge")

```

### coordinate system

- Default coordinate system in ggplot is cartesian coordinate system where x and y positions act independently


#### coordflip

- coord_flip() switches x and y axes
```{r}

ggplot(data=mpg,mapping=aes(x=class,y=hwy))+
  geom_boxplot()

ggplot(data=mpg,mapping=aes(x=class,y=hwy))+
  geom_boxplot()+
  coord_flip()


```

#### coord_quickmap()

- coord_quickmap() sets the aspect ratio correctly for maps
```{r}

nz <- map_data("nz")

# without setting aspect ratio
ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

# using coord_quickmap with aspect ratio
ggplot(data=nz,aes(long,lat,group=group))+
  geom_polygon(fill="white",color="black")+
  coord_quickmap()

```

#### coord_polar()

- coord_polar() uses polar coordinates. Polar coordinates reveal an interesting connection between a bar chart and a Coxcomb chart.


```{r}
bar <- ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=cut),show.legend = FALSE,
           width=1)+
  theme(aspect.ratio = 1)+
  labs(x=NULL,y=NULL)

bar+coord_flip()
bar+coord_polar()
```
