---
title: "Wrangling R cheat sheet"
author: "Shanmukh"
date: "April 8, 2019"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    df_print: paged
---

```{r echo=FALSE, warning=FALSE, message = FALSE}
library(stringr) # to manipulate strings
library(htmlwidgets) # to activate html widgets
library(tidyverse)# for data wrangling
library(readr)# to load files
```

# Tibble

- A tibble, or tbl_df, is a modern reimagining of the data.frame

- Tibbles are printer friendly, which makes them easier to use with large datasets containing complex objects.

- Tibble shows first 10 rows and all columns fit to the screen

- Tibble never changes the type of the inputs (e.g. it never converts strings to factors!)

- Tibble never changes the names of variables

* syntax to convert dataframe to tibble

  - `as_tibble(dataframe)` converts dataframe to tibble

* syntax to convert tibble to dataframe
  - `as.data.frame(tibble)` converts tibble to dataframe

**dataframe**
```{r }
class(iris) # verify class of iris dataset
iris_tbl<-as_tibble(iris) # Converting dataframe to tibble data set
class(iris_tbl) # verify class of tibble dataset
glimpse(iris_tbl) # provide structure of tibble similar to str(dataframe)
```

# Data Transformation

tidyr package of R is used to organizes data into tidy format

* There are two fundamental verbs of data tidying:
    - gather() transforms a wide table into a long table
    - spread() transforms a long table into a wide table

tidyr also provides separate() and extract() functions which makes it easier to pull apart a column that represents multiple variables. The complement to separate() is unite().



**messy dataframe**
```{r echo=FALSE, warning=FALSE, message = FALSE}
library(tidyr) # load tidyr package

# Creation of messy_dataset with name, drug (a and b), and heart rate
messy <- data.frame( name = c("Wilbur", "Petunia", "Gregory"), 
a = c(67, 80, 64), 
b = c(56, 90, 50) )

messy

```

### gather()
- transforms a wide table into a long table

```{r}
# gather function used to move column a & b in to drug and all values to column heart rate
# Option1
messy %>% gather(`a`,`b`,key=drug,value=hear_rate)

# Option2

#messy %>% gather(key =drug,value =hear_rate,-name)

```
**messy_2 dataframe**
```{r echo=FALSE}
set.seed(10) 

messy_2 <- data.frame( id = 1:4, 

trt = sample(rep(c('control', 'treatment'), each = 2)), 

work.T1 = runif(4), 

home.T1 = runif(4), 

work.T2 = runif(4), 

home.T2 = runif(4) )

messy_2
```

```{r}
# gather function is used to move work.T1, home.T1, work.T, home.T2 in to a column location and all values to time
messy2_g<-messy_2 %>%
  gather(key=location,value = time,-id,-trt)

# view messy2_g dataframe
messy2_g
```
**messy3 dataframe**
```{r echo=FALSE}

messy3 <- data.frame( name = rep(c('Afghanistan'), each = 4), 
year = rep(c(1999 , 2000), each = 2), 
type = c('cases','population','cases','population'),
count = c(745,19987071,2666,20595360)
)
messy3
```

## spread
- transforms a long table into a wide table
```{r}
# spread function used to move cases and population to separate columns and count column will be values for the cases and population column
messy3 %>%
  spread(key =type, value=count)

```
## separate

- seprate function teases variables apart by using splitting patterns

```{r}
# separate function used to split location in to location and ti with separater .
messy2_g %>% 
    separate(location, into = c("location", "ti"), sep = "\\.")
```

  

# Data Manipulation

* The dplyr package contains five key data manipulation functions, also called verbs:
    - select() returns a subset of the columns,
    - filter() returns a subset of the rows,
    - arrange() that reorders the rows according to single or multiple variables,
    - mutate()  used to add columns from existing data,
    - summarize() which reduces each group to a single row by calculating aggregate measures.

**billboard tibble**
```{r echo=FALSE}
billboard <- read.csv("https://raw.githubusercontent.com/shapinna/Codeporfolio/master/Data/billboard.csv", header=TRUE)

billboard_tbl <- as_tibble(billboard)

billboard_tbl

```

## select()

- Returns subset of columns

```{r}
# select only few columns from the tibble.

billboard_tbl %>%
  select(year,artist.inverted,track,time,genre)


```
## filter() 
- returns a subset of the rows

```{r warning=FALSE}
billboard_rock <-billboard_tbl %>%
filter(genre=="Rock")

billboard_rock

```
## arrange() 
- reorders the rows according to single or multiple variables
```{r}
billboard %>%
arrange(desc(date.entered))
```
## mutate()
- used to add columns from existing data
```{r}
billboard %>%
  separate(time, into=c("hours","minutes"),sep=":",convert = TRUE) %>%
  mutate(new_mins=hours*60)

```
**who** dataset  
```{r}
head(who) # top five records of who dataset

```

```{r}
# gather transforms the wide data to long data
# mutate change the values of the types for newrel to new_rel
# arrange sorts the data based on cases
# separate the type data in to three columns col1, method and sexage

new_who <- who %>%
  gather(key="Types",value = "Cases",new_sp_m014:newrel_f65,na.rm = TRUE)%>%
  mutate(Types=stringr::str_replace(Types, "newrel","new_rel"))%>%
  arrange(desc(Cases))%>%
  separate(Types,into=c("Col1","Method","SexAge"),sep="_")

new_who
```


## summarize()
- which reduces each group to a single row by calculating aggregate
```{r}

new_who_group <- dplyr::group_by(new_who,Method)  # groups the data by method
dplyr::summarize(new_who_group,count=sum(Cases))  # summarizes the data by method sum of cases

```

**Referred from tidyr overview**
[Link](https://tidyr.tidyverse.org/)
